---
title: "Martin's File"
author: "Martin Olarte"
date: "2022-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(haven)
library(dplyr)
library(stringr)
library(purrr)
library(tidyverse)
library(MASS)
GSS2018 <- read_sav("GSS2018.sav")
View(GSS2018)
```

```{r}
n <- ncol(GSS2018)
labels_vector <- map_chr(1:n, function(x) attr(GSS2018[[x]], "label"))
variable_name <- names(GSS2018)
code_book <- data.frame(variable_name, description = labels_vector)

relig_codebook <- code_book[grep("relig|strength of affiliation|god|jew|denomination|fundamentalist|bible|muslim|belief|church|pray", labels_vector, ignore.case = TRUE),]

code_book[grep("happ", labels_vector, ignore.case = TRUE),]

relig_df <- GSS2018 %>% 
  select(relig_codebook$variable_name)

View(relig_df)


all_vars <- code_book[grep("relig|strength of affiliation|god|jew|denomination|fundamentalist|bible|muslim|belief|church|pray|Confidence in scientific community", labels_vector, ignore.case = TRUE),]

all_df <- GSS2018 %>% 
  select(all_vars$variable_name)

View(all_df)
```

```{r}
least_na <- relig_df %>% 
  summarise(across(everything(), ~ sum(is.na(.)))) %>% 
  pivot_longer(cols = c(1:82)) %>% 
  arrange(value) %>%
  filter(value < nrow(relig_df) * 0.1)

the_df <- all_df %>%
  dplyr::select(least_na$name, CONSCI)

## fit ordered logit model and store results 'm'
m <- polr(as.factor(CONSCI) ~ ., data = the_df, Hess=TRUE)

## view a summary of the model
summary(m)

ci <- confint(m) # default method gives profiled CIs

exp(cbind(OR = coef(m), ci))

#odds of being less confident in science decreases 13% for every one unit decrease in frequency prayed


# getting probs from predicted


```

```{r}
library(lmreg)

religions <- data.frame(binaries(as.factor(all_df$RELIG)))
colnames(religions) <- c("CHRISTIAN", "CATHOLIC", "NONE", "PROTESTANT", "MOSLEM/ISLAM", "ORTHODOX-CHRISTIAN", "BUDDHISM", "JEWISH", "OTHER", "HINDUISM", "NA_RELIGION", "INTER-NONDENOMINATIONAL", "OTHER EASTERN", "NATIVE AMERICAN")

new_df <- the_df %>%
  cbind(religions) %>%
  dplyr::select(-c("MOSLEM/ISLAM", "ORTHODOX-CHRISTIAN", "BUDDHISM", "JEWISH", "OTHER", "HINDUISM", "NA_RELIGION", "INTER-NONDENOMINATIONAL", "OTHER EASTERN", "NATIVE AMERICAN"))

new_model <- polr(as.factor(CONSCI) ~ ., data = new_df, Hess=TRUE)

## view a summary of the model
summary(new_model)

ctable <- coef(summary(new_model))
## calculate and store p values
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2

## combined table
ctable <- cbind(ctable, "p_value" = p)

a = data.frame(ctable) %>%
  mutate(significance = case_when(
    p_value < 0.1 ~ "*",
    p_value < 0.05 ~ "**",
    p_value < 0.01 ~ "***",
    TRUE ~ ""
    ))
signif <- a %>% 
  filter(significance != "") %>%
  rownames()
  
  
ci <- confint(new_model) # default method gives profiled CIs

ratios_table <- data.frame(exp(cbind(OR = coef(new_model), ci)))
ratios_table %>%
  filter(row.names(.) %in% signif)

#for every one unit decrease in frequency prayed, the odds of being more confident in science increases 13% 
#for every one unit increase in attendance to religious services, the odds of being more confident in science increases 6% 
#for every one unit decrease in level of devotion to the Bible, the odds of being more confident in science increases 41% 
#for every one unit decrease in how fundamentalist the respondent was, the odds of being more confident in science increases 20% 

```

We can also get confidence intervals for the parameter estimates. These can be obtained either by profiling the likelihood function or by using the standard errors and assuming a normal distribution. Note that profiled CIs are not symmetric (although they are usually close to symmetric). If the 95% CI does not cross 0, the parameter estimate is statistically significant.

The CIs for PRAY, BIBLE, and FUND16 do not include 0; ATTEND does. The estimates in the output are given in units of ordered logits, or ordered log odds. The coefficients from the model can be somewhat difficult to interpret because they are scaled in terms of logs. Thus, we can interpret logistic regression models by converting the coefficients into odds ratios (OR). To get the OR and confidence intervals, we just exponentiate the estimates and confidence intervals.

- for every one unit decrease in frequency prayed, the odds of being more confident in science increases 13%
- for every one unit increase in attendance to religious services, the odds of being more confident in science increases 6% 
- for every one unit decrease in level of devotion to the Bible, the odds of being more confident in science increases 41% 
- for every one unit decrease in how fundamentalist the respondent was, the odds of being more confident in science increases 20% 


```{r}
relevel_df <- new_df %>%
  mutate(CONSCI_re = case_when(
    CONSCI == 1 ~ 3,
    CONSCI == 3 ~ 1,
    CONSCI == 2 ~ 2,
    TRUE ~ NA_real_
  )) %>%
  dplyr::select(-CONSCI)

new_model <- polr(as.factor(CONSCI_re) ~ ., data = relevel_df, Hess=TRUE)

## view a summary of the model
summary(new_model)

ctable <- coef(summary(new_model))
## calculate and store p values
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2

## combined table
ctable <- cbind(ctable, "p_value" = p)

a = data.frame(ctable) %>%
  mutate(significance = case_when(
    p_value < 0.1 ~ "*",
    p_value < 0.05 ~ "**",
    p_value < 0.01 ~ "***",
    TRUE ~ ""
    ))
signif <- a %>% 
  filter(significance != "") %>%
  rownames()
  
  
ci <- confint(new_model) # default method gives profiled CIs

ratios_table <- data.frame(exp(cbind(OR = coef(new_model), ci)))
ratios_table %>%
  filter(row.names(.) %in% signif)
```

- for every one unit decrease in frequency prayed, the odds of being more confident in science increases 13%
- for every one unit increase in attendance to religious services, the odds of being more confident in science increases 6% 
- for every one unit decrease in level of devotion to the Bible, the odds of being more confident in science increases 70% 
- for every one unit decrease in how fundamentalist the respondent was, the odds of being more confident in science increases 24% 
